---
title: 'Assignment: Course Project 1'
author: "Asylbek Zhainazarov"
date: "May 2, 2016"
output: md_document
keep_md: yes
---

##Loading and preprocessing the data

Loading the data

```{r, echo=TRUE}
activity <- read.table("activity.csv", sep = ",", header = TRUE)
head(activity)
summary(activity)
```

##What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day


```{r, echo=TRUE}
tapply(activity$step, activity$date, sum,na.rm = TRUE)
```

2. Make a histogram of the total number of steps taken each day

```{r, echo=TRUE}
hist(tapply(activity$step, activity$date, sum,na.rm = TRUE),main="Histogram of the total number of steps taken each day",xlab="Number of steps")
```

3. Calculate and report the mean and median of the total number of steps taken per day

- The mean of the total number of steps taken per day


```{r, echo=TRUE}
x <- tapply(activity$step, activity$date, sum,na.rm = TRUE)
x <- as.vector(x)
mean(x)
```

- The median of the total number of steps taken per day

```{r, echo=TRUE}
median(x)
```

##What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
g <- split(activity$steps,activity$interval)
plot(sapply(g,mean,na.rm = TRUE),type = "l",main="Time-serious plot of average number of steps",ylab = "Average number of steps",xlab="Index of 5-minute interval")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE}
f <- as.numeric(sapply(g,mean,na.rm = TRUE))
max(f)
imax = 0
for (i in 1:length(f)) {
  if (f[i] == max(f)){
    imax = i
    break
  }
}
imax
```
##Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=TRUE}
sapply(activity, function(x) sum(is.na(x)))
```
The total number of missing values in the dataset:

```{r, echo=TRUE}
row.has.na <- apply(activity, 1, function(x){any(is.na(x))})
sum(row.has.na)
```

2. Devise a strategy for filling in all of the missing values in the dataset

We will use the mean for attribute in the entire dataset to fill the missing values for any given day.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in


```{r, echo=TRUE}
activitynew <- activity # a new dataset
y <- tapply(activitynew$step, activitynew$date, mean,na.rm = TRUE)
activitynew$steps[which(is.na(activitynew$steps))] <- mean(y[is.finite(y)])
```

The total number of missing values in the new dataset:

```{r, echo=TRUE}
row.has.na <- apply(activitynew, 1, function(x){any(is.na(x))})
sum(row.has.na)
```

4. Make a histogram of the total number of steps taken each day 

```{r, echo=TRUE}
hist(tapply(activity$step, activitynew$date, sum),main="Histogram of the total number of steps taken each day",xlab="Number of steps")
```
Calculate and report the mean and median total number of steps taken per day. 

```{r, echo=TRUE}
x <- tapply(activitynew$step, activitynew$date, sum)
x <- as.vector(x)
mean(x)
```

- The median of the total number of steps taken per day

```{r, echo=TRUE}
median(x)
```

Do these values differ from the estimates from the first part of the assignment?

Yes, these values differ from the estimates from the first part of the assignment.

What is the impact of imputing missing data on the estimates of the total daily number of steps?

Now the mean and median of the total number of steps taken per day are equal.

##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
activitynew$date <- as.Date(activitynew$date)
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activitynew$wDay <- c('weekend', 'weekday')[(weekdays(activitynew$date) %in% weekdays1)+1L]
head(activitynew)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 


```{r, echo=TRUE}
z <- split(activitynew, activitynew$wDay)
names(z)
head(z$weekday)
head(z$weekend)

g <- split(z$weekday$steps,z$weekday$interval)
plot(sapply(g,mean),type = "l",main="Weekday",ylab = "Average number of steps",xlab="Index of 5-minute interval")

g1 <- split(z$weekend$steps,z$weekend$interval)
plot(sapply(g1,mean),type = "l",main="Weekend",ylab = "Average number of steps",xlab="Index of 5-minute interval")
```